---
title: <center><br>Introduction to ggplot2</br></center>
author: <center>Davis Vaughan</center>
date: <center>September 18, 2016</center>
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: True
---
<center>

No matter what level of R Programmer you are, you likely struggle to remember every single intricate detail required by ggplot2. It has a steep learning curve compared to base R's plotting functions, but once you learn to use it, it can produce production level graphics. This post will serve as a quick reference guide to ggplot2. The idea is that it can quickly give an example of the code and output for each type of plot you might need.

---

#Scatter Plots {.tabset .tabset-fade}

I often need scatter plots to show relationships between data. Let's look at a quick example of the benefits of ggplot vs base R.

##Base R
```{r}
plot(cars)
```

---

##ggplot2
```{r}
library(ggplot2)
library(plotly,warn.conflicts = FALSE)
ggplotly(qplot(data=cars,x=speed,y=dist,geom="point"),height = 400, width = 600)
```

---

#Line Graphs


</center>