---
title: <center><br>Introduction to ggplot2</br></center>
author: <center>Davis Vaughan</center>
date: <center>September 18, 2016</center>
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: True
---
<center>

```{r include=FALSE}
#Libraries
library(ggplot2)
library(plotly, warn.conflicts = FALSE)
```

No matter what level of R Programmer you are, you likely struggle to remember every single intricate detail required by ggplot2. It has a steep learning curve compared to base R's plotting functions, but once you learn to use it, it can produce production level graphics. This post will serve as a quick reference guide to ggplot2. The idea is that it can quickly give an example of the code and output for each type of plot you might need.

---

#Some Motivation {.tabset .tabset-fade}

One might say, "R already makes nice graphics!" So, before we get started, let's quickly compare ggplot vs base R with a simple scatter plot example. As you can see below, the plot created using a combination of ggplot's quick plot function and plotly is clearly superior over base R.

##Base R
```{r}
plot(cars)
```

---

##ggplot2
```{r}
#Create a nice plot using "quick plot"
prettyPlot <- qplot(data=cars, x=speed, y=dist, geom="point")
#Make the plot interactive using ggplotly
ggplotly(prettyPlot, height = 400, width = 600)
```

---

#Scatter Plots

It's time to get into the nitty gritty. Let's learn to make a scatter plot using ggplot2. Here is the basic syntax.

```{r scatterExample,fig.height=4, fig.width=6 }
ggplot(data=cars) +
  geom_line( aes(x = speed, y = dist) )
```

Let's break this down. It seems like a lot just to make a scatter plot, and it is, but let's give ggplot2 the benefit of the doubt. What it lacks in wordiness, it makes up for in flexibility. Here is what is happening:

<div align="left">

1) ggplot(data=cars) - The basic function for setting up the plot. Here, I commonly define the data.frame that I will be pulling my data from using the data paramater. Cars is just a built in dataset.

2) "+" - Yes, the plus sign is very important. One way to think about this is "adding" a geometry to your plot (like a line). [Here](http://stackoverflow.com/questions/38166708/plus-sign-between-ggplot2-and-other-function-r) is something that might clear this up.

3) geom_line( aes(x = speed, y = dist) ) - This function has quite a few parts.
    * geom - This stands for "geometry," and is a common way to add elements to a ggplot graph. A common geometry is a line, like what is shown here. 
    * aes() - This is the aesthetics function. This is where you add your x and y components. It can also add a number of other things, like color (or colour as Hadley the creator of ggplot2 would say). This can actually help with creating a legend, so check that section for more information. Also, see [this](http://stackoverflow.com/questions/11657380/is-there-a-table-or-catalog-of-aesthetics-for-ggplot2) stackoverflow post for a complete list of default aesthetics that you can play with.

</div>

---

#Line Graphs

---

#Legends

---

#Quick Plots with qplot()

---

#Interactive Plots With Plotly

